---
author: "Alec L. Robitaille"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Homework: Week 8
2021-09-08 [updated: `r Sys.Date()`]

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1

### Data
```{r}
DT <- data_reedfrogs()
precis(DT)
```


```{r}
writeLines(readLines(tar_read(stan_d_file_w08_model_frogs_1)))
```

```{r}
model_frogs_1_draws <- tar_read(stan_d_mcmc_w08_model_frogs_1)$draws()

mcmc_areas(model_frogs_1_draws, regex_pars = 'alpha')
mcmc_areas(model_frogs_1_draws, regex_pars = 'p\\[')
```

```{r}
writeLines(readLines(tar_read(stan_d_file_w08_model_frogs_2)))
```

```{r}
model_frogs_2_draws <- tar_read(stan_d_mcmc_w08_model_frogs_2)$draws()
mcmc_areas(model_frogs_2_draws, regex_pars = 'predation')
```

```{r}
writeLines(readLines(tar_read(stan_d_file_w08_model_frogs_3)))
```

```{r}
model_frogs_3_draws <- tar_read(stan_d_mcmc_w08_model_frogs_3)$draws()
mcmc_areas(model_frogs_3_draws, regex_pars = 'size')
```


```{r}
writeLines(readLines(tar_read(stan_d_file_w08_model_frogs_4)))
```

```{r}
model_frogs_4_draws <- tar_read(stan_d_mcmc_w08_model_frogs_4)$draws()

mcmc_areas(model_frogs_4_draws, regex_pars = 'beta')
```

Negative influence of predation is somewhat balanced by size of tank. 


<!-- TODO use AIC -->


## Question 2

> In 1980, a typical Bengali woman could have 5 or more children in her
lifetime. By the year 2000, a typical Bengali woman had only 2 or 3. You’re
going to look at a historical set of data, when contraception was widely
available but many families chose not to use it. These data reside in
`data(bangladesh)` and come from the 1988 Bangladesh Fertility Survey. Each row
is one of 1934 women. There are six variables, but you can focus on two of them
for this practice problem: (1) district: ID number of administrative district
each woman resided in (2) use.contraception: An indicator (0/1) of whether the
woman was using contraception. 

```{r}
DT <- data_bangladesh()
precis(DT)
```

Now, focus on predicting `use.contraception`, clustered by `district_id`. Fit
both (1) a traditional fixed-effects model that uses an index variable for
district and (2) a multilevel model with varying intercepts for district. Plot
the predicted proportions of women in each district using contraception, for
both the fixed-effects model and the varying-effects model. That is, make a plot
in which district ID is on the horizontal axis and expected proportion using
contraception is on the vertical.Make one plot for each model, or layer them on
the same plot, as you prefer. How do the models disagree? Can you explain the
pattern of disagreement? In particular, can you explain the most extreme cases
of disagreement, both why they happen where they do and why the models reach
different inferences?


Fixed effects model

```{r}
writeLines(readLines(tar_read(stan_d_file_w08_model_bang_fixed)))
```

```{r}
model_bang_fixed_draws <- tar_read(stan_d_mcmc_w08_model_bang_fixed)$draws()

mcmc_areas(model_bang_fixed_draws, regex_pars = 'alpha', transformations = inv.logit)
```

Multilevel  model

```{r}
writeLines(readLines(tar_read(stan_d_file_w08_model_bang_multi)))
```

```{r}
model_bang_multi_draws <- tar_read(stan_d_mcmc_w08_model_bang_multi)$draws()

mcmc_trace(model_bang_multi_draws)
mcmc_areas(model_bang_multi_draws, regex_pars = 'alpha', transformations = inv.logit)
mcmc_areas(model_bang_multi_draws, pars = c('alpha_bar', 'sigma'))
```


Comparison

```{r}
(mcmc_areas(model_bang_fixed_draws, regex_pars = 'alpha', transformations = inv.logit) + labs(title = 'fixed')) + 
	(mcmc_areas(model_bang_multi_draws, regex_pars = 'alpha\\[', transformations = inv.logit) + labs(title = 'multilevel'))

setDT(model_bang_fixed_draws)
setDT(model_bang_multi_draws)

compare <- rbindlist(list(
	melt(model_bang_fixed_draws, measure.vars = patterns('alpha'))[, model_type := 'fixed'],
	melt(model_bang_multi_draws, measure.vars = patterns('alpha'))[, model_type := 'multilevel']
), fill = TRUE)

compare[, variable := as.integer(gsub('alpha\\[|\\]', '', variable))]
ggplot(compare[, .(value = mean(value)), .(variable, model_type)]) + 
	geom_hline(yintercept = 0, alpha = 0.5) + 
	geom_point(aes(variable, value, color = model_type)) +
	geom_line(aes(variable, value, group = model_type), alpha = 0.2) +
	scale_color_viridis_d(begin = 0.3, end = 0.8) + 
	labs(x = 'district', y = 'alpha')
```



## Question 3

> Return to the Trolley data, `data(Trolley)`, from Chapter 12. Define and fit a
varying intercepts model for these data. By this I mean to add an intercept
parameter for the individual to the linear model. Cluster the varying intercepts
on individual participants, as indicated by the unique values in the id
variable. Include action, intention, and contact as before. Compare the varying
intercepts model and a model that ignores individuals, using both WAIC/LOO and
posterior predictions. What is the impact of individual variation in these data?

```{r}
DT <- data_trolley()
precis(DT)

model_data <- c(
	as.list(DT[, .(response, action, intention, contact)]), 
	N = DT[, .N],
	K = DT[, uniqueN(response) - 1]
)
```


```{cmdstan, output.var = 'model_trolley', cache = TRUE}
data {
  int N;
  int K;
  int response[N];
  int action[N];
  int intention[N];
  int contact[N];
}
parameters {
	// Cut points are the positions of responses along cumulative odds
  ordered[K] cutpoints;
  real beta_action;
  real beta_intention;
  real beta_contact;
}
transformed parameters {
  vector[N] phi;

	for (i in 1:N) {
		phi[i] = beta_action * action[i] + beta_contact * contact[i] + beta_intention * intention[i];
	}
}
model {
	response ~ ordered_logistic(phi, cutpoints);
  cutpoints ~ normal(0, 1.5);
  beta_action ~ normal(0, 0.5);
  beta_contact ~ normal(0, 0.5);
  beta_intention ~ normal(0, 0.5);
}
generated quantities {
	vector[N] log_lik;
  for (i in 1:N) {
    log_lik[i] = ordered_logistic_lpmf(response[i] | phi[i], cutpoints);
  }
}

```


```{r, cache = TRUE}
# model_trolley_stan <- 'model_trolley.stan'
# writeLines(readLines(model_trolley_stan))
# model_trolley <- cmdstan_model(model_trolley_stan, cpp_options = list(stan_threads = TRUE))

model_trolley_sample <-
	model_trolley$sample(
		data = model_data,
		chains = 4,
		save_warmup = FALSE,
		parallel_chains = 4,
		threads_per_chain = 4
	)
model_trolley_draws <- cmd_draws(model_trolley_sample)

mcmc_areas(model_trolley_draws, regex_pars = 'cutpoints')
mcmc_areas(model_trolley_draws, regex_pars = 'beta')

model_trolley_loo <- model_trolley_sample$loo()
```


```{r}
model_data <- c(
	as.list(DT[, .(response, action, intention, contact, individual)]), 
	N = DT[, .N],
	N_individual = DT[, uniqueN(individual)],
	K = DT[, uniqueN(response) - 1]
)
```

```{cmdstan, output.var = 'model_trolley_multi', cache = TRUE}
data {
  int N;
  int K;
  int response[N];
  int action[N];
  int intention[N];
  int contact[N];
  int individual[N];
  int N_individual;
}
parameters {
	// Cut points are the positions of responses along cumulative odds
  ordered[K] cutpoints;
  real beta_action;
  real beta_intention;
  real beta_contact;
	real alpha_bar;
	real sigma;
  vector[N_individual] alpha;
}
transformed parameters {
  vector[N] phi;

	for (i in 1:N) {
		phi[i] = alpha[individual[i]] + beta_action * action[i] + beta_contact * contact[i] + beta_intention * intention[i];
	}
}
model {
	// Hyper parameter priors
	alpha_bar ~ normal(0, 1.5);
	sigma ~ exponential(1);

	// Priors
	alpha ~ normal(alpha_bar, sigma);

	response ~ ordered_logistic(phi, cutpoints);
  cutpoints ~ normal(0, 1.5);
  beta_action ~ normal(0, 0.5);
  beta_contact ~ normal(0, 0.5);
  beta_intention ~ normal(0, 0.5);
}
generated quantities {
	vector[N] log_lik;
  for (i in 1:N) {
    log_lik[i] = ordered_logistic_lpmf(response[i] | phi[i], cutpoints);
  }
}

```

```{r, cache = TRUE}
# model_trolley_multi_stan <- 'model_trolley_multi.stan'
# writeLines(readLines(model_trolley_multi_stan))
# model_trolley_multi <- cmdstan_model(model_trolley_multi_stan, cpp_options = list(stan_threads = TRUE))

model_trolley_multi_sample <-
	model_trolley_multi$sample(
		data = model_data,
		chains = 4,
		save_warmup = FALSE,
		parallel_chains = 4,
		threads_per_chain = 4
	)
model_trolley_multi_draws <- cmd_draws(model_trolley_multi_sample)

mcmc_areas(model_trolley_multi_draws, regex_pars = 'cutpoints')
mcmc_areas(model_trolley_multi_draws, regex_pars = 'beta')

model_trolley_multi_loo <- model_trolley_multi_sample$loo()
```

```{r}
col_patterns <- '^beta|^alpha|^sigma|^cutpoints'
setDT(model_trolley_draws)
setDT(model_trolley_multi_draws)

precis(model_trolley_draws[, .SD, .SDcols = patterns(col_patterns)])
precis(model_trolley_multi_draws[, .SD, .SDcols = patterns(col_patterns)])

compared <- loo_compare(model_trolley_loo, model_trolley_multi_loo)
print(compared, simplify = FALSE)
```
