---
title: "Week 2"
author: "Alec L. Robitaille"
date: "2021-08-24 [updated: `r Sys.Date()`]"
output: github_document
---

\newcommand{\ols}[1]{\mskip.5\thinmuskip\overline{\mskip-.5\thinmuskip {#1} \mskip-.5\thinmuskip}\mskip.5\thinmuskip} % overline short


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
```


## Question 1
The weights listed below were recorded in the !Kung census, but heights were not
recorded for these individuals. Provide predicted heights and 89% compatibility
intervals for each of these individuals. That is, fill in the table below, using
model-based predictions.

```
Individual, weight, expected height, 89% interval
1, 45,,,
2, 40,,,
3, 65,,,
4, 31,,,
5, 53,,,
```

Model:

$h_{i} \sim \text{Normal}(\mu_{i}, \sigma)$

$\mu_{i} = \alpha + \beta(x_{i} - \bar{x})$

$\alpha \sim \text{Normal}(178, 20)$

$\beta \sim \text{Log-Normal}(0, 1)$

$\sigma \sim \text{Uniform}(0, 50)$

```{r}
library(rethinking)
library(data.table)
library(ggplot2)

data(Howell1)

d <- Howell1[Howell1$age >= 18,]

m <- quap(
	alist(
		height ~ dnorm(mu, sigma),
		mu <- a + b * (weight - mean(d$weight)),
		a ~ dnorm(178, 20),
		b ~ dnorm(0, 1),
		sigma ~ dunif(0, 50)
	), 
	data = d
)

```
