---
title: "Week 1"
author: "Alec L. Robitaille"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Variables

N: fixed by experimenter

p: Prior probability

W: A probability distribution given the data. 

## Joint model

W ~ Binomial(N , p)

p ~ Uniform(0, 1)

W is distributed binomially with N observations and probability p on each

p is distributed uniformally at 1

## Question 1

Suppose the globe tossing data (Chapter 2) had turned out to be 4 water
in 15 tosses. Construct the posterior distribution, using grid approximation.
Use the same flat prior as in the book.


```{r}
# Size of grid for grid approximation
gridsize <- 1000

# Prior grid
prior_grid <- seq(0, 1, length.out = gridsize)

# Prior probability (all 1)
prior_prob <- rep(1, gridsize)

# Data probability
#  given 4/15, using binomial distribution
data_prob <- dbinom(4, 15, prob = vals)

# Calculate the posterior numerator by multiplying prior and data probability
posterior_num <- prior_prob * data_prob
# Standardize by sum of posterior numerator
posterior <- posterior_num / sum(posterior_num)

plot(posterior)
```

