---
title: "Week 2"
author: "Alec L. Robitaille"
date: "2021-08-24 [updated: `r Sys.Date()`]"
output: github_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
```


## Question 1
The weights listed below were recorded in the !Kung census, but heights were not
recorded for these individuals. Provide predicted heights and 89% compatibility
intervals for each of these individuals. That is, fill in the table below, using
model-based predictions.

```
Individual, weight, expected height, 89% interval
1, 45,,,
2, 40,,,
3, 65,,,
4, 31,,,
5, 53,,,
```

Model:

$h_{i} \sim \text{Normal}(\mu_{i}, \sigma)$

$\mu_{i} = \alpha + \beta(x_{i} - \bar{x})$

$\alpha \sim \text{Normal}(178, 20)$

$\beta \sim \text{Log-Normal}(0, 1)$

$\sigma \sim \text{Uniform}(0, 50)$

```{r}
library(rethinking)
library(data.table)
library(ggplot2)
library(tidybayes)

theme_set(theme_bw())

data(Howell1)

d <- Howell1[Howell1$age >= 18,]

m <- quap(
	alist(
		height ~ dnorm(mu, sigma),
		mu <- a + b * (weight - mean(d$weight)),
		a ~ dnorm(178, 20),
		b ~ dnorm(0, 1),
		sigma ~ dunif(0, 50)
	), 
	data = d
)

precis(m)
```

Simulate:

```{r sim}
# Set weights to simulate for
weights <- data.table(weight = c(45, 40, 65, 31, 54),
											id = as.character(seq(1, 5)))
simmed <- sim(m, list(weight = weights$weight), n = 1e3)

# Tidy
DT <- melt(as.data.table(simmed), measure.vars = paste0('V', 1:5),
					 value.name = 'height', variable.name = 'id')
DT[, id := gsub('V', '', id)]
DT[weights, weight := weight, on = 'id']

# Plot
ggplot(DT, aes(height)) +
	stat_halfeye(.width = .89) + 
	facet_wrap(~id)
```


## Question 2

Model the relationship between height (cm) and the natural logarithm of weight
(log-kg): log(weight). Use the entire Howell1 data frame, all 544 rows, adults
and non-adults. Use any model type from Chapter 4 that you think useful: an
ordinary linear regression, a polynomial or a spline. Plot the posterior
predictions against the raw data

```{r}
library(rethinking)
library(data.table)
library(ggplot2)
library(tidybayes)

theme_set(theme_bw())

data(Howell1)
d <- Howell1
d$logweight <- log(d$weight)

m1 <- quap(
	alist(
		height ~ dnorm(mu, sigma),
		mu <- a + b * (logweight - mean(d$logweight)),
		a ~ dnorm(178, 20),
		b ~ dnorm(0, 1),
		sigma ~ dunif(0, 50)
	), 
	data = d
)
sim_x <- log(1:60)
simmed <- sim(m1, list(logweight = sim_x))

# Tidy
DT <- melt(as.data.table(simmed), value.name = 'height', variable.name = 'x')
DT[data.table(sim_x, x = paste0('V', 1:60)),
	 logweight := sim_x,
	 on = 'x']
DT[, meanheight := mean(height), by = logweight]
DT[, low := PI(height)[1], by = logweight]
DT[, high := PI(height)[2], by = logweight]

# Plot
ggplot(DT) + 
	geom_ribbon(aes(x = exp(logweight), ymin = low, ymax = high), fill = 'grey') + 
	geom_point(aes(exp(logweight), height), data = d, color = 'lightblue', alpha = 0.8) + 
	geom_line(aes(exp(logweight), meanheight))
```

Using the dlnorm, prior of a Log normal distribution on beta

```{r}
m2 <- quap(
	alist(
		height ~ dnorm(mu, sigma),
		mu <- a + b * (logweight - mean(d$logweight)),
		a ~ dnorm(178, 20),
		b ~ dlnorm(0, 1),
		sigma ~ dunif(0, 50)
	), 
	data = d
)
sim_x <- log(1:60)
simmed <- sim(m2, list(logweight = sim_x))

# Tidy
DT <- melt(as.data.table(simmed), value.name = 'height', variable.name = 'x')
DT[data.table(sim_x, x = paste0('V', 1:60)),
	 logweight := sim_x,
	 on = 'x']
DT[, meanheight := mean(height), by = logweight]
DT[, low := PI(height)[1], by = logweight]
DT[, high := PI(height)[2], by = logweight]

# Plot
ggplot(DT) + 
	geom_ribbon(aes(x = exp(logweight), ymin = low, ymax = high), fill = 'grey') + 
	geom_point(aes(exp(logweight), height), data = d, color = 'lightblue', alpha = 0.8) + 
	geom_line(aes(exp(logweight), meanheight))
```



## Question 3

Set up:

```{r}
library(rethinking)
library(data.table)
library(ggplot2)
library(tidybayes)

theme_set(theme_bw())

data(Howell1)
d <- Howell1
d$weight_s <- scale(d$weight)
d$weight_s2 <- scale(d$weight) ^ 2

m <- quap(
	alist(
		height ~ dnorm(mu, sigma),
		mu ~ a + b1 * weight_s + b2 * weight_s2,
		a ~ dnorm(178, 20),
		b1 ~ dlnorm(0, 1),
		b2 ~ dnorm(0, 1),
		sigma ~ dunif(0, 50)
	),
	data = d
)


n <- 20
sim_x <- seq(min(d$weight_s), max(d$weight_s), length.out = n)
linked <- link(
	m,
	data = list(weight_s = sim_x, weight_s2 = sim_x ^ 2),
	post = extract.prior(m)
)[1:50,]
plot(NULL, xlim = range(sim_x), ylim = range(linked) + c(-10, 10))
apply(linked, 1, FUN = function(x) lines(sim_x, x))
```

