---
title: "Week 2"
author: "Alec L. Robitaille"
date: "2021-08-24 [updated: `r Sys.Date()`]"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
```


## Question 1
The weights listed below were recorded in the !Kung census, but heights were not
recorded for these individuals. Provide predicted heights and 89% compatibility
intervals for each of these individuals. That is, fill in the table below, using
model-based predictions.

```
Individual, weight, expected height, 89% interval
1, 45,,,
2, 40,,,
3, 65,,,
4, 31,,,
5, 53,,,
```

Model:

$h_{i} \sim \text{Normal}(\mu_{i}, \sigma)$

$\mu_{i} = \alpha + \beta(x_{i} - \bar{x})$

$\alpha \sim \text{Normal}(178, 20)$

$\beta \sim \text{Log-Normal}(0, 1)$

$\sigma \sim \text{Uniform}(0, 50)$

```{r}
library(rethinking)
library(data.table)
library(ggplot2)
library(tidybayes)

theme_set(theme_bw())

data(Howell1)

d <- Howell1[Howell1$age >= 18,]

m <- quap(
	alist(
		height ~ dnorm(mu, sigma),
		mu <- a + b * (weight - mean(d$weight)),
		a ~ dnorm(178, 20),
		b ~ dnorm(0, 1),
		sigma ~ dunif(0, 50)
	), 
	data = d
)

precis(m)
```

Simulate:

```{r sim}
# Set weights to simulate for
weights <- data.table(weight = c(45, 40, 65, 31, 54),
											id = as.character(seq(1, 5)))
simmed <- sim(m, list(weight = weights$weight), n = 1e3)

# Tidy
DT <- melt(as.data.table(simmed), measure.vars = paste0('V', 1:5),
					 value.name = 'height', variable.name = 'id')
DT[, id := gsub('V', '', id)]
# (Not needed with stat eye width)
# DT[, low := PI(height)[1], by = id]
# DT[, high := PI(height)[2], by = id]
DT[weights, weight := weight, on = 'id']

# Plot
ggplot(DT, aes(height)) +
	stat_halfeye(.width = .89) + 
	facet_wrap(~id)
```


## Question 2

Model the relationship between height (cm) and the natural logarithm of weight
(log-kg): log(weight). Use the entire Howell1 data frame, all 544 rows, adults
and non-adults. Use any model type from Chapter 4 that you think useful: an
ordinary linear regression, a polynomial or a spline. Plot the posterior
predictions against the raw data

```{r}

```
