---
author: "Alec L. Robitaille"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Week 3 
2021-08-25 [updated: `r Sys.Date()`]

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

All three problems below are based on the same data. The data in `data(foxes)`
are 116 foxes from 30 different urban groups in England. These foxes are like
street gangs. Group size varies from 2 to 8 individuals. Each group maintains
its own (almost exclusive) urban territory. Some territories are larger than
others. The area variable encodes this information. Some territories also have
more avgfood than others. We want to model the weight of each fox. For the
problems below, assume this DAG


### Setup

```{r}
# DAG
dag <- dagify(
	weight ~ groupsize + avgfood,
	groupsize ~ avgfood,
	avgfood ~ area,
	exposure = 'area',
	outcome = 'weight'
)

dag_plot(dag)
```



## Question 1
> Use a model to infer the total causal influence of area on weight. Would
increasing the area available to each fox make it heavier (healthier)? You might
want to standardize the variables. Regardless, use prior predictive simulation
to show that your model’s prior predictions stay within the possible outcome
range.


### Workings

Area on weight  

scale(weight) ~ dnorm(mu, sigma)  
mu <- a + b * (scale(area))  
a: intercept  
when weight and area are scaled, the expected intercept is 0  
therefore  
a ~ dnorm(0, 0.5)  

b: beta, rate of change given one unit of increase in area  
b ~ dnorm(0, 1)  

sigma: standard deviation  
uniform prior  
sigma ~ dunif(0, 50)  

### Model

```{r}
data(foxes)

foxes$scale_area <- scale(foxes$area)
foxes$scale_weight <- scale(foxes$weight)

m1 <- quap(
	alist(
		scale_weight ~ dnorm(mu, sigma),
		mu <- a + bArea * scale_area,
		a ~ dnorm(0, 0.05),
		bArea ~ dnorm(0, 0.5),
		sigma ~ dunif(0, 50)
	), 
	data = foxes
)
```

### Prior predictive simulation

```{r}
prior <- extract.prior(m1)
l <- link(m1, post = prior, data = list(scale_area = c(-2, 2)))
plot_link(l, 20)
```

### Paths
Interest: Area on Weight

Paths

1. Area -> Avgfood -> Weight
2. Area -> Avgfood -> Groupsize -> Weight

Avgfood and Groupsize are pipes between Area and Weight. There are 
no backdoors or colliders. 

### Interpretation

> Would increasing the area available to each fox make it heavier (healthier)? 

`bArea` has a mean of 0.02, with compatibility intervals around 0. 
Therefore the model does not indicate a total causal influence of 
area on the weight. 


```{r}
precis(m1)

post <- extract.samples(m1)
s <- sim(m1, data = list(scale_area = c(-2, 2)), post = post)

ggplot(data.table(s), aes(V1)) +
	stat_halfeye(.width = .89)
```


## Question 2
> Now infer the causal impact of adding food to a territory. Would this make foxes
heavier? Which covariates do you need to adjust for to estimate the total causal
influence of food?


### Paths
Interest: Food on weight

Paths: 

1. Food -> Weight
2. Food -> Groupsize -> Weight

Groupsize is a pipe between Area and Weight. There are no backdoors or colliders. 


### Model
```{r}
foxes$scale_avgfood <- scale(foxes$avgfood)

m2 <- quap(
	alist(
		scale_weight ~ dnorm(mu, sigma),
		mu <- a + bFood * scale_avgfood,
		a ~ dnorm(0, 0.05),
		bFood ~ dnorm(0, 0.5),
		sigma ~ dunif(0, 50)
	), 
	data = foxes
)

precis(m2)

post <- extract.samples(m2)
s <- sim(m2, data = list(scale_avgfood = c(-2, 2)), post = post)

ggplot(data.table(s), aes(V1)) +
	stat_halfeye(.width = .89)
```


### Interpretation

> Would this make foxes heavier? Which covariates do you need to adjust for to
estimate the total causal influence of food?

`bFood` has a mean of -0.02, with compatibility intervals around 0. 
The model does not indicate a total causal influence of area on the weight. 
No covariates are needed to estiamte the total causal influence of food 
because there are no backdoors or colliders. 


## Question 3

> Now infer the causal impact of group size. Which covariates do you need to
adjust for? Looking at the posterior distribution of the resulting model, what
do you think explains these data? That is, can you explain the estimates for all
three problems? How do they go together?


### Paths
Interest: Group size on weight

Paths: 

1. Groupsize -> Weight
2. Groupsize <- Avgfood -> Weight

Avgfood is a collider between Groupsize and Weight. There is a backdoor
on Groupsize and the path is closed. 

```{r}
foxes$scale_groupsize <- scale(foxes$groupsize)

m3 <- quap(
	alist(
		scale_weight ~ dnorm(mu, sigma),
		mu <- a + bGroupsize * scale_groupsize + bFood * scale_avgfood,
		a ~ dnorm(0, 0.05),
		bGroupsize ~ dnorm(0, 0.5),
		bFood ~ dnorm(0, 0.5),
		sigma ~ dunif(0, 50)
	), 
	data = foxes
)

precis(m3)

DT <- melt(data.table(extract.samples(m3))[, .(bGroupsize, bFood)])
ggplot(DT) + 
	geom_density(aes(value, fill = variable), alpha = 0.6) + 
	theme_bw() + scale_fill_viridis_d(begin = 0.3, end = 0.8)
```

### Interpretation

> Which covariates do you need to adjust for? Looking at the posterior
distribution of the resulting model, what do you think explains these data? That
is, can you explain the estimates for all three problems? How do they go
together?

The Avgfood covariate needs to be included since it is a collider between
Groupsize and Weight. The mean and compatibility intervals of bFood are positive,
while the mean and compatibility intervals of bGroupsize are negative. This
indicates food's positive relationship with weight could be buffered or interacting
with the negative relationship of group size. Increased food leads to increased
body weight, but more food also results in larger groups, which decreases the 
food availability. 
