---
author: "Alec L. Robitaille"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Homework: Week 6
2021-09-06 [updated: `r Sys.Date()`]

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
options(digits = 2, scipen = 999)
```

### Setup

```{r}
# Packages
library(ggdag)
library(dagitty)
library(data.table)
library(ggplot2)
library(rethinking)
library(cmdstanr)
library(posterior)
library(bayesplot)

# Functions
dag_plot <- function(dag) {
	stat <- node_status(dag, FALSE)
	stat$data$status[is.na(stat$data$status)] <- 'intermediate'
	ggplot(stat, aes(x = x, y = y, xend = xend, yend = yend)) +
	  geom_dag_point(aes(color = status), alpha = 0.5, size = 15) +
	  geom_dag_edges() +
		labs(color = '') + 
	  geom_dag_text(color = 'black') +
		scale_color_manual(values = list('exposure' = '#35608DFF',
																		 'outcome' = '#22A884FF',
																		 'intermediate' = 'grey50')) + 
	  theme_void()
}

cmd_draws <- function(model) {
	as_draws_df(model$draws())
}
```

## Question 1

> The data in `data(NWOGrants)` are outcomes for scientific funding applications
for the Netherlands Organization for Scientific Research (NWO) from 2010–2012
(see van der Lee and Ellemers doi:10.1073/pnas.1510159112). These data have a
very similar structure to the UCBAdmit data discussed in Chapter 11. I want you
to consider a similar question: What are the total and indirect causal effects
of gender on grant awards? Consider a mediation path (a pipe) through dis-
cipline. Draw the corresponding DAG and then use one or more binomial GLMs to
answer the question.


## Data

```{r}
data(NWOGrants)

precis(NWOGrants)
summary(NWOGrants)
```


```{r}
register_knitr_engine(override = FALSE)
```

```{cmdstan, output.var = 'q1_prior'}
```
