---
author: "Alec L. Robitaille"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Week 7
2021-09-07 [updated: `r Sys.Date()`]

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Data

```{r}
DT <- data_trolley()
precis(DT)
```

Response: 1-7 integer, "how morally permissible the action to be taken (or not) is". Categorical, ordered, but distances between categories is not metric or known. 

Logit = log-odds, cumulative logit = log-cumulative-odds. Both constrained between 0-1. 

Log-cumulative-odds for response 7 will be infinity since log(1/(1-1)) = infinity. 
Given this, we only need K-1 = 6 intercepts. 


### Model: ordered categorical outcome

Probability of data: $R_{i} \sim \text{Ordered-logit}(\phi_{i}, K)$

Linear model: $\phi_{i} = 0$

Prior for each intercept: $K_{k} \sim \text{Normal}(0, 1.5)$


```{r}
writeLines(readLines(tar_read(stan_c_file_w07_model1)))
```


```{r}
model1_draws <- tar_read(stan_c_mcmc_w07_model1)$draws()

mcmc_areas(model1_draws, regex_pars = 'beta')
```

Interactions

```{r}
writeLines(readLines(tar_read(stan_c_file_w07_model1_interactions)))
```

```{r}
model1_interactions_draws <- tar_read(stan_c_mcmc_w07_model1_interactions)$draws()

mcmc_areas(model1_interactions_draws, regex_pars = 'beta')
mcmc_areas(model1_interactions_draws, regex_pars = 'cut', transformations = inv.logit) + xlim(0, 1)
```


### Model: ordered categorical predictors

Probability of data: $R_{i} \sim \text{Ordered-logit}(\phi_{i}, K)$

Linear model: $\phi_{i} = \beta_{E} \sum_{j=0}^{E_{i}-1}\delta_{j} + \beta{A_{i}}A_{i} + \beta{I_{i}}I_{i} + \beta{C_{i}}C_{i}$

Prior for each intercept: $K_{k} \sim \text{Normal}(0, 1.5)$
Prior for each $\beta$: $\beta_{A}, \beta_{I}, \beta_{C}, \beta_{E}K_{k} \sim \text{Normal}(0, 1)$
Prior for the $\delta$ vector: $\delta \sim \text{Dirichlet}(\alpha)$

Dirichlet distribution = multivariate extension of the beta distribution. Probabilities between zero and one, that all sum to one. It is parameterized by
pseudo-counts of observations. 

The intercept takes the first category, so we add 0 to the sequence of

Make sure to reorder education levels 

```{r}
writeLines(readLines(tar_read(stan_c_file_w07_model2)))
```

```{r}
model2_draws <- tar_read(stan_c_mcmc_w07_model2)$draws()

mcmc_areas(model2_draws, regex_pars = 'beta')
mcmc_areas(model2_draws, regex_pars = 'delta')
mcmc_areas(model2_draws, regex_pars = 'cut', transformations = inv.logit) + xlim(0, 1)
```


## Question 1

> In the Trolley data — `data(Trolley)` — we saw how education level (modeled as an
ordered category) is associated with responses. Is this association causal? One
plausible confound is that education is also associated with age, through a
causal process: People are older when they finish school than when they begin
it. Reconsider the Trolley data in this light. Draw a DAG that represents
hypothetical causal relationships among response, education, and age. 


```{r}
dag <- dagify(
  response ~ education + age + action + intention + contact,
  education ~ age,
  contact ~ action,
  exposure = 'education',
  outcome = 'response'
)

dag_plot(dag)
adjustmentSets(dag, exposure = 'education', outcome = 'response', effect = 'total' )
```


Which statistical model or models do you need to evaluate the causal influence of
education on responses? Fit these models to the trolley data. What do you
conclude about the causal relationships among these three variables?


```{r}
writeLines(readLines(tar_read(stan_c_file_w07_model2_age)))
```

```{r}
model2_age_draws <- tar_read(stan_c_mcmc_w07_model2_age)$draws()

mcmc_areas(model2_age_draws, regex_pars = 'beta')
mcmc_areas(model2_age_draws, regex_pars = 'delta')
mcmc_areas(model2_age_draws, regex_pars = 'cut', transformations = inv.logit) + xlim(0, 1)
```


## Question 2

> Consider one more variable in the Trolley data: Gender. Suppose that gender
might influence education as well as response directly. Draw the DAG now that
includes response, education, age, and gender. Using only the DAG, is it
possible that the inferences from Problem 1 are con founded by gender? If so,
define any additional models you need to infer the causal influence of education
on response. What do you conclude?

```{r}
dag <- dagify(
  response ~ education + age + gender + action + intention + contact,
  education ~ age,
  education ~ gender,
  contact ~ action,
  exposure = 'education',
  outcome = 'response'
)

dag_plot(dag)
adjustmentSets(dag, exposure = 'education', outcome = 'response', effect = 'total' )
```


```{r}
writeLines(readLines(tar_read(stan_c_file_w07_model2_gender)))
```

```{r}
model2_gender_draws <- tar_read(stan_c_mcmc_w07_model2_gender)$draws()

mcmc_areas(model2_gender_draws, regex_pars = 'beta')
mcmc_areas(model2_gender_draws, regex_pars = 'delta')
mcmc_areas(model2_gender_draws, regex_pars = 'cut', transformations = inv.logit) + xlim(0, 1)
```
